---
title: "Comprehensive Examination Responses"
author: "Jake Thompson"
date: "`r Sys.Date()`"
bibliography: comps_ref.bib
biblio-style: apalike
csl: apa.csl
link-citations: yes
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE,
  out.width = "80%",
  fig.align = 'center',
  fig.show = "hold"
)
```

## Introduction

This document is a compressed version of a more detailed document available [online](https://wjakethompson.github.io/soccer/). Specifically, this document aims to answers the specific questions for my written comprehensive examination. As such, this document may reference parts of the more complete document, but will also be comprehensive in containing all information needed to adequately address the questions.

## Gather data

The first step to this analysis is to gather the data for the English Premier League. For this analysis, I will scrape the data from [ESPN](http://www.espnfc.us/). I will first define a function that can scrape the team names and webpages from the league page. This functions uses the **purrr** [@R-purrr] and **rvest** [@R-rvest] packages to scrape the information, and **dplyr** [@R-dplyr] to format the output.

```{r league-func, message = FALSE}
library(dplyr)
library(purrr)
library(rvest)

scrape_league <- function(x) {
  cont <- TRUE
  while(cont) {
    url_data <- safe_read_html(x)
    
    if(is.null(url_data[[1]])) {
      closeAllConnections()
      Sys.sleep(5)
    } else {
      url_data <- url_data[[1]]
      cont <- FALSE
    }
  }
  
  league_table <- url_data %>%
    html_nodes(css = "table") %>%
    html_table()
  league_table <- league_table[[1]]
  colnames(league_table) <- as.character(league_table[1,])
  colnames(league_table) <- make.names(colnames(league_table), unique = TRUE)
  league_table <- league_table[-1,]
  
  league_table <- league_table %>%
    select(club = TEAM, goals_for = `F`, goals_against = A, points = PTS) %>%
    mutate(club = trimws(club, which = "both"))
  
  teams <- url_data %>%
    html_nodes("td a") %>%
    html_text() %>%
    as.character() %>%
    trimws(which = "both")
  team_urls <- url_data %>%
    html_nodes("td a") %>%
    html_attr("href") %>%
    as.character()
  
  league_table <- league_table %>%
    left_join(data_frame(club = teams, club_url = team_urls), by = "club") %>%
    as_data_frame()
  
  return(league_table)
}
```

We can then scrape use the `scrape_league` function to get the team names and URLs.

```{r scrape-teams}
safe_read_html <- safely(read_html)
epl <- scrape_league("http://www.espnfc.us/english-premier-league/23/table")
epl
```

I then define a function for scraping the game information from each team's webpage.

```{r game-func}
scrape_team <- function(x, y) {
    x <- gsub("/index", "/fixtures", x, fixed = TRUE)
    
    cont <- TRUE
    while(cont) {
      url_data <- safe_read_html(x)
      
      if(is.null(url_data[[1]])) {
        closeAllConnections()
        Sys.sleep(5)
      } else {
        url_data <- url_data[[1]]
        cont <- FALSE
      }
    }
    date <- url_data %>%
      html_nodes(".headline") %>%
      html_text() %>%
      as.character()
    if ("LIVE" %in% date) {
      date[which(date == "LIVE")] <- format(Sys.Date(), "%b %d, %Y")
    }
    date <- mdy(date)
    home_team <- url_data %>%
      html_nodes(".score-home-team .team-name") %>%
      html_text() %>%
      as.character()
    away_team <- url_data %>%
      html_nodes(".score-away-team .team-name") %>%
      html_text() %>%
      as.character()
    home_score <- url_data %>%
      html_nodes(".home-score") %>%
      html_text() %>%
      as.character() %>%
      gsub(" ", "", x = .) %>%
      gsub( " *\\(.*?\\) *", "", x = .) %>%
      as.numeric()
    away_score <- url_data %>%
      html_nodes(".away-score") %>%
      html_text() %>%
      as.character() %>%
      gsub(" ", "", x = .) %>%
      gsub( " *\\(.*?\\) *", "", x = .) %>%
      as.numeric()
    competition <- url_data %>%
      html_nodes(".score-column.score-competition") %>%
      html_text() %>%
      as.character()
    
    team_data <- data_frame(
      date = date,
      home = home_team,
      away = away_team,
      home_goals = home_score,
      away_goals = away_score,
      competition = competition
    ) %>%
      arrange(date) %>%
      unique()
    
    abbrev <- as_data_frame(table(c(team_data$home, team_data$away))) %>%
      top_n(n = 1, wt = n) %>%
      select(Var1) %>%
      flatten_chr()
    
    if (nrow(team_data) < 3) {
      ret_data <- data_frame(
        club = y,
        abbrev = y,
        team_data = NA
      )
    } else {
      if (abbrev == "Sporting") {
        team_data$home[which(team_data$home == "Sporting")] <- y
        team_data$away[which(team_data$away == "Sporting")] <- y
        ret_data <- data_frame(
          club = y,
          abbrev = y,
          team_data = list(team_data)
        )
      } else {
        team_data <- filter(team_data, home != "Sporting", away != "Sporting")
        ret_data <- data_frame(
          club = y,
          abbrev = abbrev,
          team_data = list(team_data)
        )
      }
    }
    
    return(ret_data)
}
```

And then I use that function to scrape game data. After the game data is scraped, I filter to only include games within the Premier League, and do some cleaning (e.g., replace ESPN abbreviations with the real club name).

```{r scrape-games, message = FALSE}
library(lubridate)
epl_games <- map2_df(.x = epl$club_url, .y = epl$club, .f = scrape_team)

team_lookup <- select(epl_games, -team_data)

epl_games <- bind_rows(epl_games$team_data) %>%
  unique() %>%
  arrange(date, home) %>%
  left_join(select(epl_games, -team_data), by = c("home" = "abbrev")) %>%
  rename(home_club = club) %>%
  left_join(select(epl_games, -team_data), by = c("away" = "abbrev")) %>%
  rename(away_club = club) %>%
  mutate(
    real_home = ifelse(is.na(home_club), home, home_club),
    real_away = ifelse(is.na(away_club), away, away_club),
    home = real_home,
    away = real_away
  ) %>%
  select(-(home_club:real_away)) %>%
  filter(!(date < Sys.Date() & is.na(home_goals))) %>%
  filter(date > ymd("2016-03-01")) %>%
  rename(h_goals = home_goals, a_goals = away_goals) %>%
  filter(home %in% epl$club, away %in% epl$club, competition == "Prem",
    !is.na(h_goals))

knitr::kable(head(epl_games), caption = "English Premier League Games")
```

## Estimate the models

I estimate the bivariate Poisson and the game random intercept model using Stan [@R-rstan; @stan]. 


## References
